The Multi-Stencil Language

### 2016-05-27 ARTICLES/IJPP/article/article.tex
IJPP article biblio

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}
Contributions :
\begin{itemize}
\item Formalism of a multi-stencil program and its parallelization
\item The Multi-stencil Language: agnostic from the type of mesh and of the choosen implementation
\item A first implementation of the MSL compiler (static scheduling + fusion + dump to SkelGIS/components)
\item Evaluation of the different types of parallelism introduced + compiler + fusion
\end{itemize}
%------------------------------------------------------------------------------
\section{Related work}
\label{sect:rel}
\input{related.tex}
%------------------------------------------------------------------------------
\section{Computational model of Multi-Stencil Programs}
\label{sect:formalism}
\input{formalism.tex}
%------------------------------------------------------------------------------
\section{The Multi-Stencil Language}
\label{sect:msl}
\input{msl.tex}
%------------------------------------------------------------------------------
\section{Parallelization of Multi-Stencil Programs}
\label{sect:parallelism}
\input{parallelism.tex}
%------------------------------------------------------------------------------
\section{Static Scheduling and performance model}
\label{sect:msp}
\input{msp.tex}
%------------------------------------------------------------------------------
\section{Implementation}
\label{sect:comp}
\input{compiler.tex}
%------------------------------------------------------------------------------
\section{Evaluation}
\label{sect:eval}
\input{evaluation.tex}
%------------------------------------------------------------------------------
\section{Conclusion}
\label{sect:concl}
%------------------------------------------------------------------------------

>>>>

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}
\input{introduction.tex}
%------------------------------------------------------------------------------
\section{Related work}
\label{sect:rel}
\input{related.tex}
%------------------------------------------------------------------------------
\section{Computational model of Multi-Stencil Programs}
\label{sect:formalism}
\input{formalism.tex}
%------------------------------------------------------------------------------
\section{The Multi-Stencil Language}
\label{sect:msl}
\input{msl.tex}
%------------------------------------------------------------------------------
\section{Parallelization of Multi-Stencil Programs}
\label{sect:parallelism}
\input{parallelism.tex}
%------------------------------------------------------------------------------
\section{Static Scheduling and performance model}
\label{sect:msp}
\input{msp.tex}
%------------------------------------------------------------------------------
\section{Implementation}
\label{sect:comp}
\input{compiler.tex}
%------------------------------------------------------------------------------
\section{Evaluation}
\label{sect:eval}
\input{evaluation.tex}
%------------------------------------------------------------------------------
\section{Conclusion}
\label{sect:concl}
%------------------------------------------------------------------------------

### 2016-05-27 ARTICLES/IJPP/article/compiler.tex
IJPP article biblio

SkelGIS is a C++ embedded language~\cite{} which offers distributed data structures for numerical simulations, and programming interfaces to easily write codes using them. The code produced by SkelGIS uses MPI (message Passing Interface~\cite{}). Thus, this code can be used on distributed memory architectures such as clusters of machines. The distributed data structures offered by SkelGIS are, first, a distributed two dimensional Cartesian mesh~\cite{}, and second, a distributed graph of Cartesian meshes (hybrid mesh)~\cite{}. SkelGIS is used in this paper for the dump of MSL and more precisely to produce the code associated to the data parallelization technique.

OpenMP~\cite{} is a language based on directives added to C/C++ or Fortran. Those directives offers a way to take advantage of shared memory architectures such as multi-processors, multi-cores, many-cores or even GPUs (in OpenMP 4.0). Among other available directives, some of them offer the possibility to manage a pool of threads and to produce a task parallel program. In this paper OpenMP is used in its version 3.0, which means that we do not take advantage of the dynamic schedulers available for OpenMP 4.0 such as libgomp~\cite{}, StarPU~\cite{}, or XKaapi~\cite{}. Those dynamic schedulers will be the subject of future work. In this paper, however, is evaluated the performance of the static scheduling introduced in Section~\ref{sect:msp}.

The dump step of the MSL compiler automatically generates an hybrid code which uses the SkelGIS Cartesian distributed data structure and an OpenMP static scheduling generated from the algorithm previously described in Section~\ref{sect:tsp}. To facilitate the maintainability of this hybrid code, a component-based programming model is used, the Low Level Component Model~\cite{}.

@@

One can notice that MSL could use different languages as it is mesh-agnostic. For example, a future work is to use the distributed unstructured mesh proposed by PamPA~\cite{}. Moreover, MSL could use pthreads instead of OpenMP to manage the static scheduling.

>>>>

SkelGIS is a C++ embedded language~\cite{CPE:CPE3494} which offers distributed data structures for numerical simulations, and programming interfaces to easily write codes using them. The code produced by SkelGIS uses MPI (message Passing Interface~\cite{Graham2009MSE}). Thus, this code can be used on distributed memory architectures such as clusters of machines. The distributed data structures offered by SkelGIS are, first, a distributed two dimensional Cartesian mesh~\cite{DBLP:conf/ieeehpcs/HeleneS13}, and second, a distributed graph of Cartesian meshes (hybrid mesh)~\cite{DBLP:conf/europar/CoullonL14}. SkelGIS is used in this paper for the dump of MSL and more precisely to produce the code associated to the data parallelization technique.

OpenMP~\cite{660313} is a language based on directives added to C/C++ or Fortran. Those directives offers a way to take advantage of shared memory architectures such as multi-processors, multi-cores, many-cores or even GPUs (in OpenMP 4.0). Among other available directives, some of them offer the possibility to manage a pool of threads and to produce a task parallel program. In this paper OpenMP is used in its version 3.0, which means that we do not take advantage of the dynamic schedulers available for OpenMP 4.0 such as libgomp~\footnote{\url{https://gcc.gnu.org/projects/gomp/}}, StarPU~\cite{Augonnet2011}, or XKaapi~\cite{Gautier:2013:XRS:2510661.2511383}. Those dynamic schedulers will be the subject of future work. In this paper, however, is evaluated the performance of the static scheduling introduced in Section~\ref{sect:msp}.

The dump step of the MSL compiler automatically generates an hybrid code which uses the SkelGIS Cartesian distributed data structure and an OpenMP static scheduling generated from the algorithm previously described in Section~\ref{sect:tsp}. To facilitate the maintainability of this hybrid code, a component-based programming model is used, the Low Level Component Model~\cite{l2c} (\llc).

@@

One can notice that MSL could use different languages as it is mesh-agnostic. For example, a future work is to use the distributed unstructured mesh proposed by PaMPA~\cite{lachat:hal-00768916}. Moreover, MSL could use pthreads instead of OpenMP to manage the static scheduling.

### 2016-05-27 ARTICLES/IJPP/article/evaluation.tex
IJPP article biblio

The first parallelization technique proposed by MSL is a data parallelization. To implement this parallelization, the dump step of the MSL compiler uses the SkelGIS language. This language, as already described, offers a distributed Cartesian mesh and programming interfaces to use it as a sequential data structure. It has been evaluated itself compared to an MPI implementation on the same FullSWOF2D simulation~\cite{}. What MSL brings to SkelGIS is an automatic detection of where MPI synchronizations are needed during the numerical simulation, and how to make a fusion of some computations in order to reduce cache misses. This section gives a performance comparison of the code produced by MSL for FullSWOF2D, and of the equivalent manual SkelGIS program (with manual synchronizations and fusion choices). Moreover, as the code produced by MSL is a component-based code, this evaluation also shows that no overheads are introduced \llc~\cite{}. Following evaluations have been performed onto the machine of the TGCC Curie described in Table~\ref{tab:TGCC}. Each evaluation has been performed ten times and the median is presented in results.

>>>>

The first parallelization technique proposed by MSL is a data parallelization. To implement this parallelization, the dump step of the MSL compiler uses the SkelGIS language. This language, as already described, offers a distributed Cartesian mesh and programming interfaces to use it as a sequential data structure. It has been evaluated itself compared to an MPI implementation on the same FullSWOF2D simulation~\cite{CPE:CPE3494}. What MSL brings to SkelGIS is an automatic detection of where MPI synchronizations are needed during the numerical simulation, and how to make a fusion of some computations in order to reduce cache misses. This section gives a performance comparison of the code produced by MSL for FullSWOF2D, and of the equivalent manual SkelGIS program (with manual synchronizations and fusion choices). Moreover, as the code produced by MSL is a component-based code, this evaluation also shows that no overheads are introduced \llc~\cite{l2c}. Following evaluations have been performed onto the machine of the TGCC Curie described in Table~\ref{tab:TGCC}. Each evaluation has been performed ten times and the median is presented in results.
